Simulation Study: AdaPT with null model option (step covariate)

### 2018-09-04 datasets/simulations/simulations-informative-cosine-nullAdaPT.Rmd
Add supplementary analyses of AdaPT with null option

---
title: "Simulation Study: Informative Covariate (cosine)"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

@@

## intermediary files we create below
gauss_file <- file.path(resdir, "informative-cosine-benchmark-gaussian.rds")
tdist_file <- file.path(resdir, "informative-cosine-benchmark-t5.rds")
tdist11_file <- file.path(resdir, "informative-cosine-benchmark-t11.rds")
chisq_file <- file.path(resdir, "informative-cosine-benchmark-chisq4.rds")

@@

@@

```{r gauss-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

@@

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

@@

```{r gauss-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
```

@@

```{r gauss-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r gauss-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

@@

@@

```{r t5-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

@@

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

@@

```{r t5-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
```

@@

```{r t5-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r t5-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

@@

@@

Next, we consider a second setting with t-distributed noise.

@@

```{r t11-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

@@

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

@@

```{r t11-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
```

@@

```{r t11-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r t11-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

@@

@@

```{r}
bdchi <- modifyBMethod(bd, "ashq", mode = "empirical")
```

@@

```{r chisq4-one-simulation}
onerun <- simIteration(bdchi, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

@@

Since `ashq` is not appropriate when the test statistics are not approximately normal or t,
we also plot the results without this method.

```{r chisq4-metrics-averages-noash, results = "hide"}
plotsim_average(resdf, met = "rejections", filter_set = c(excludeSet, "ashq"),
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met = "FDR", filter_set = c(excludeSet, "ashq"),
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met = "TPR", filter_set = c(excludeSet, "ashq"),
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met = "TNR", filter_set = c(excludeSet, "ashq"),
                merge_ihw = TRUE, errorBars = TRUE) 
```

@@

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods. Again, without `ashq`.

```{r chisq4-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = c(excludeSet, "ashq")) >= 3) {
    aggupset(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
             alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

@@

>>>>

---
title: "Simulation Study: AdaPT with null model option (cosine covariate)" 
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

This set of simulations differs from `simulations-informative-cosine.Rmd` only
in the implementation of the AdaPT method for multiple testing correction. This
is a follow up analysis similar to that described and performed in
`simulations-informative-step-nullAdaPT.Rmd`. For more details on the parameter settings
of the two additional AdaPT methods included in the comparison, please check the
other `-nullAdaPT.Rmd` analysis.

The analysis in `simulations-informative-step-nullAdaPT.Rmd` was carried out
to determine whether the additional `adapt-withnull` and `adapt-onlynull` methods
could provide an improvement over the `adapt-glm` parameter setting in the case
of low informativeness in the independent covariate. This analysis aims to examine
how the `adapt-withnull` and `adalpt-onlynull` parameter settings perform in
comparison to the default `adapt-glm` settings when the covariate is more
informative.

@@

## intermediary files we create below
gauss_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-gaussian.rds")
tdist_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-t5.rds")
tdist11_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-t11.rds")
chisq_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-chisq4.rds")

@@

We add a couple alternative calls to AdaPT to test whether the method performance improves in cases
of low "informativeness" if given the option to not use the informative covariate. This option is
coded by the null model, `"pi0 ~ 1"`, where only an intercept term is included in the model. Specifically,
we add two alternative calls to the benchmark - one call with the null model added as one of the candidate
models to `pi_formulas=` and `mu_formulas=`, and a second call with the null model given as the only
candidate model to `pi_formulas=` and `mu_formulas=`.

```{r}
bdplus <- addBMethod(bdplus, "adapt-withnull",
                     adaptMT::adapt_glm,
                     function(x) { x$qvals },
                     pvals = pval,
                     x = data.frame(icov = ind_covariate),
                     pi_formulas = c("~ 1", paste0("splines::ns(icov, df = ", seq(2, 10, 2), ")")),
                     mu_formulas = c("~ 1", paste0("splines::ns(icov, df = ", seq(2, 10, 2), ")")),
                     alphas = 0)

bdplus <- addBMethod(bdplus, "adapt-onlynull",
                     adaptMT::adapt_glm,
                     function(x) { x$qvals },
                     pvals = pval,
                     x = data.frame(icov = ind_covariate),
                     pi_formulas = "~ 1",
                     mu_formulas = "~ 1",
                     alphas = 0)

candycols2 <- rbind(candycols,
                    data.frame(Method = c("adapt-withnull", "adapt-onlynull"),
                               col = c("darkgoldenrod3", "darkgoldenrod3"),
                               lty = c("dashed", "dotted"),
                               stringsAsFactors = FALSE))
```

@@

```{r gauss-one-simulation}
onerun <- simIteration(1, bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

@@

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, 
                palette = candycols2)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, 
                palette = candycols2)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, 
                palette = candycols2)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r gauss-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

@@

```{r gauss-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
                  palette = candycols2)

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
                  palette = candycols2)
```

@@

```{r gauss-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
                  palette = candycols2)
```

@@

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

Again, we recreate the plots after subsetting on AdaPT methods.

```{r gauss-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

@@

```{r t5-one-simulation}
onerun <- simIteration(1, bdplust, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

@@

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r t5-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

@@

```{r t5-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
                  palette = candycols2)

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
                  palette = candycols2)
```

@@

```{r t5-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
                  palette = candycols2)
```

@@

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

Again, we recreate the plots after subsetting on AdaPT methods.

```{r t5-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

@@

@@

```{r t11-one-simulation}
onerun <- simIteration(1, bdplust, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

@@

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r t11-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

@@

```{r t11-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
                  palette = candycols2)

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
                  palette = candycols2)
```

@@

```{r t11-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
                  palette = candycols2)
```

@@

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

Again, we recreate the plots after subsetting on AdaPT methods.

```{r t11-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

@@

```{r}
bdchi <- modifyBMethod(bdplus, "ashq", mode = "empirical")
```

We also drop the FDRreg methods from the chi-squared setting.

```{r}
bdchi <- dropBMethod(bdchi, "fdrreg-t")
bdchi <- dropBMethod(bdchi, "fdrreg-e")
```

@@

```{r chisq4-one-simulation}
onerun <- simIteration(1, bdchi, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

@@

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r chisq4-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

@@

@@

Again, we recreate the plots after subsetting on AdaPT methods.

```{r chisq4-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

### 2018-09-04 datasets/simulations/simulations-informative-step-nullAdaPT.Rmd
Add supplementary analyses of AdaPT with null option

---
title: "Simulation Study: AdaPT with null model option (cosine covariate)" 
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

In this set of simulations, we consider settings with both null and non-null
tests with an informative covariate. The covariate is sampled uniformly from
the interval [0, 1], and the conditional probability of a test being non-null
is a smooth (cosine) function of the covariate. We include simulation results
again with Gaussian, t-distributed, and Chi-Squared distributed noise.

This set of simulations differs from `simulations-informative-cosine.Rmd` only
in the implementation of the AdaPT method for multiple testing correction. This
is a follow up analysis similar to that described and performed in
`simulations-informative-step-nullAdaPT.Rmd`. For more details on the parameter settings
of the two additional AdaPT methods included in the comparison, please check the
other `-nullAdaPT.Rmd` analysis.

The analysis in `simulations-informative-step-nullAdaPT.Rmd` was carried out
to determine whether the additional `adapt-withnull` and `adapt-onlynull` methods
could provide an improvement over the `adapt-glm` parameter setting in the case
of low informativeness in the independent covariate. This analysis aims to examine
how the `adapt-withnull` and `adalpt-onlynull` parameter settings perform in
comparison to the default `adapt-glm` settings when the covariate is more
informative.

@@

## intermediary files we create below
gauss_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-gaussian.rds")
tdist_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-t5.rds")
tdist11_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-t11.rds")
chisq_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-chisq4.rds")

@@

```{r parameters-shared}
m <- 20000                        # integer: number of hypothesis tests
pi0 <- pi0_cosine(0.90)           # numeric: proportion of null hypotheses
icovariate <- runif               # functional: independent covariate
```

>>>>

---
title: "Simulation Study: AdaPT with null model option (step covariate)" 
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

In this set of simulations, we consider settings with both null and non-null
tests with an informative covariate. The covariate is sampled uniformly from
the interval [0, 1], and the conditional probability of a test being non-null
is a non-smooth (monotone) step function of the covariate. We include simulation
results again with Gaussian, t-distributed, and Chi-Squared distributed noise.

This set of simulations differs from `simulations-informative-step.Rmd` only
in the implementation of the AdaPT method for multiple testing correction.
In the primary set of simulation results (`simulations-informative-step.Rmd`),
we observed that the AdaPT method can lose power when the informative covariate
is only weakly informative. While the method still controls the FDR, we observe
that the loss in power is substantial, with AdaPT achieving a lower mean TPR
across 100 replications when using a weakly informative covariate rather than a
completely uninformative (random) covariate.

In this set of simulations, we assess whether this loss of power with a weakly
informative covariate can be mitigated by allowing the AdaPT model selection
procedure to decide between the same collection of spline models used before, and
an intercept-only (null) model, which makes no use of the covariate. This approach
was suggested by the authors of the AdaPT method through personal communication.

Two additional implementations of the AdaPT model are included in this analysis:
`adapt-withnull`, the approach described above, and `adapt-onlynull`, where the method
is forced to ignore the informative covariate. The latter (`onlynull`) implementation
provides a baseline for comparing the two other AdaPT parameter settings.

@@

## intermediary files we create below
gauss_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-gaussian.rds")
tdist_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-t5.rds")
tdist11_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-t11.rds")
chisq_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-chisq4.rds")

@@

```{r parameters-shared}
m <- 20000                        # integer: number of hypothesis tests
pi0 <- pi0_step(0.90)             # numeric: proportion of null hypotheses
icovariate <- runif               # functional: independent covariate
```
